---
title: "C997 task submission"
author: "Jesse Ashby"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

##Section B - Data preparation and import into R

I downloaded the April 1, 2010 to July 1, 2018 census data from the site 
specified in the task requirements (United States Census Bureau, 2018).   The 
file was setup with one row per geographic summary level and had columns for 
population estimates for each year between 2010 and 2018.  The components of 
each year's estimate also each had their own columns. 

My goal in preparing the data for import into R was to extract only the relevant 
data and then transform it to meet the rules of tidy data (Grolemund & Wickham, 
2017).  I started by opening the file in Microsoft Excel and removing all of the
rows that were not for the state I live in (Texas). I then removed all columns 
except for the population estimate for each year.  This left me with 1 row and 9
columns.  I pivoted this row to 9 rows in 1 column.  This setup my response 
variable of population estimate, but I lacked the needed explanatory variable of
time.  I added a new column and manually entered the year of the population 
estimate.

To import the data into R, I used the read_csv function to get the data into a 
tibble data structure, useful for working with data in the tidy format.  

```{r echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, dataImport}
library(tidyverse)

pops <- read_csv("texasPopEstimates.csv")
pops
```

##Section A - Linear regression analysis

```{r echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, linearModel}
library(modelr)
require(scales)

# (Grolemund & Wickham, 2017) 
popsModel <- lm(popestimate ~ year, data = pops)
popsModel

vis <- ggplot(pops, aes(year, popestimate)) +
  geom_point() + 
  scale_y_continuous (labels = comma) + 
  ylab('population') +
  geom_smooth(method = "lm", se = FALSE) 

vis
```


##Section C - Statistical description using summary()

```{r echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, summary}
summary(popsModel)
```

##Section D - 5 year prediction for population of Texas

```{r}
next5years <- data_frame('year' = c((max(pops$year) + 1): (max(pops$year) + 5)))
next5pops <- data_frame('popestimate' = predict(popsModel, next5years, 
                                                type = "response"))
next5estimates <- data.frame('years' = next5years, 'popestimate' = next5pops)
pops <- rbind(pops, next5estimates)

vis <- ggplot(pops, aes(year, popestimate)) +
  geom_point() + 
  geom_text(aes(label=ifelse(year == max(pops$year),
                             as.character(as.integer(pops$popestimate)),'')),
            hjust=1.5) + 
  scale_y_continuous (labels = comma) + 
  ylab('population') +
  geom_smooth(method = "lm", se = FALSE) 

vis

```

##References:

Grolemund, G, and Wickham, H. "R For Data Science." R For Data Science, 
  O'Reilly, 2017, r4ds.had.co.nz/index.html.

United States Census Bureau. (2018).  National, State, and Puerto Rico 
  Commonwealth Totals Datasets: Population, population change, and estimated 
  components of population change: April 1, 2010 to July 1, 2018 [CSV file]. 
  Retrieved from https://www2.census.gov/programs-surveys/popest/datasets/2010-2018/national/totals/nst-est2018-alldata.csv

